---
title: "SOMED-geo project"
subtitle: "Geocoding SOMED data"
description: |
  Geocoding BAG data using two providers.
date: "`r Sys.Date()`"
author:
  - name: Radoslaw Panczak 
    url: https://github.com/RPanczak
    affiliation: ISPM
    affiliation_url: https://www.ispm.unibe.ch/
    orcid_id: 0000-0001-5141-683X
# bibliography: biblio.bib
output:
  distill::distill_article:
    highlight: pygments
    toc: true
    toc_depth: 1
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width=9, fig.height=7, dpi=300, out.width="900px", out.height="700px")

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

options(scipen=999)
set.seed(12345)

library(pacman) 
p_load(tidyverse, readxl, janitor, magrittr, 
       kableExtra, scales,
       sf, tmap, tmaptools)

tmap_mode("view") # makes map interactive
```

# Addresses 

```{r message=FALSE}
bag_somed_2018 <- read_xlsx("data-raw/BAG-open/2018_Flat_File_de.xlsx", skip = 1) %>% 
  remove_empty() %>% 
  clean_names() %>% 
  filter(!is.na(kanton)) %>%
  rowid_to_column("link") %>% 
  relocate(link)

write_rds(bag_somed_2018, "data/BAG-open/bag_somed_2018.Rds")
```

Raw file consists of dataset of `r number(length(bag_somed_2018$link), big.mark = ",")` institutions.  

Few example addresses, illustrating one case with no street address given (`NA`).

```{r}
bag_somed_2018 %>% 
  select(institution:kanton) %>% 
  slice(15:19) %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>% 
  column_spec(1, italic = TRUE) %>%
  row_spec(1, italic = FALSE)
```

# GeoAdmin 

## Info

Working with search service of [GeoAdmin API](http://api3.geo.admin.ch/services/sdiservices.html?highlight=geocode#search). 
It doesn't require API sign up. It does however require 'fair use'. [Comment](https://groups.google.com/g/geoadmin-api/c/T9uHlYD28Hc) from swisstopo staff defined it as staying below 2000 req/hour.

<aside>
Pausing for 2s after each query should do the job in case of working with large dataset.
</aside>

## Addresses with street names

```{r}
bag_addresses_street <- bag_somed_2018 %>% 
  select(link:ort) %>% 
  filter(!is.na(strasse)) %>% 
  mutate(address_orig = paste(strasse, ort, sep = ", "))

# length(bag_somed_2018$link)
# length(bag_addresses_street$link)
```

Using `r number(length(bag_addresses_street$link), big.mark = ",")` addresses that have some information on the name of the street. 


```{r geoadmin-proc, eval=FALSE, include=FALSE}
p_load(RCurl, jsonlite)

for (i in seq_along(bag_addresses_street$link)) {
  
  u <- paste0("https://api3.geo.admin.ch/rest/services/api/SearchServer?searchText=", 
              URLencode(bag_addresses_street$address_orig[i]), 
              "&type=locations")
  print(u)
  
  d <- getURL(u)
  
  print(d)
  
  j <- jsonlite::fromJSON(d, flatten = TRUE) 
  
  if (length(j$results) > 0) {
    
    j <- as.data.frame(j)
    
    names(j) <- gsub(pattern = "results.", replacement = "", x = names(j))
    
    j$link <- bag_addresses_street$link[i]
    
    if(i == bag_addresses_street$link[1]) {
      
      results_geoadmin <- j
      
    } else {
      
      results_geoadmin <- bind_rows(results_geoadmin, j)
      
    }
  }
  
  Sys.sleep(2)
}

results_geoadmin %<>% 
  as_tibble() %>% 
  clean_names() %>% 
  group_by(link) %>% 
  mutate(group_order = row_number(),
         group_n = n()) %>% 
  ungroup() %>% 
  relocate(link, fuzzy, group_order, group_n)

write_rds(results_geoadmin, "data/geocoding/results_geoadmin.Rds")
```

## Results

```{r}
results_geoadmin <- read_rds("data/geocoding/results_geoadmin.Rds")
```

`r number(length(unique(results_geoadmin$link)), big.mark = ",")` addresses have been geocoded. Unfortunately the dataset consists of  `r number(length(results_geoadmin$link), big.mark = ",")` records since some of the adrresses have multiple entries. 

`r number(nrow(filter(results_geoadmin, group_n == 1)), big.mark = ",")` addresses have clean, one geocode. 

`r number(nrow(filter(results_geoadmin, group_order == 1 & group_n > 1)), big.mark = ",")` addresses have between 2 or more hits. Mostly these are listing various addresses on same street locations

```{r}
results_geoadmin %>% 
  filter(link == 5) %>% 
  select(link, attrs_label) %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>% 
  column_spec(1, italic = TRUE) %>%
  row_spec(1, italic = FALSE)
```

In such cases first address (with bigger geocoding 'weight') is preserved. 

In extreme cases, where address is imprecise, up to 50 (*sic!*) geocodes can be returned, for instance in case of:

```{r}
bag_addresses_street %>% 
  select(-address_orig) %>% 
  filter(link == 596) %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>% 
  column_spec(1, italic = TRUE) %>%
  row_spec(1, italic = FALSE)
```

These addresses will be geocoded again with alternative method.

```{r}
problems <- c(77, 123, 329, 352, 596, 624, 635, 636, 646, 667, 682, 687, 703, 708, 759, 776, 777, 787, 884, 1093, 1120, 1123, 1131, 1132, 1136, 1139, 1144, 1149, 1156, 1168, 1175, 1191, 1193, 1204, 1264)

bag_addresses_street %>% 
  select(-address_orig) %>% 
  filter(link %in% problems) %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>% 
  column_spec(1, italic = TRUE) %>%
  row_spec(1, italic = FALSE)
```


```{r"}
results_geoadmin_clean <- results_geoadmin %>% 
  filter(!link %in% problems) %>% 
  group_by(link) %>% 
  filter(group_order == 1) %>% 
  select(link, attrs_detail, attrs_y, attrs_x) %>% 
  rename(address_geocoder = attrs_detail,
         geoy = attrs_y, geox = attrs_x) %>% 
  mutate(source = "geoadmin")

rm(results_geoadmin)
```

Clean dataset at this stage consists of `r number(nrow(results_geoadmin_clean), big.mark = ",")` addresses. 

```{r}
fail_geoadmin <- bag_addresses_street %>% 
  anti_join(select(results_geoadmin_clean, link))
```

`r number(nrow(fail_geoadmin), big.mark = ",")` addresses failed to be geocoded despite having addres and will be attempted with next method. 

# Google

## Info

Working via `ggmap` [package](https://github.com/dkahle/ggmap). 
It requires [API sign up](https://cran.r-project.org/web/packages/ggmap/readme/README.html) 

<aside>
... so please be careful with reruns of large batches of data!
</aside>

```{r google-auth, eval=FALSE, include=FALSE}
p_load(ggmap)

key_google <- readr::read_file("secrets/api_google.txt")
register_google(key = key_google)
```

## Addresses with incomplete info 

```{r}
bag_addresses_no_street <- bag_somed_2018 %>% 
  select(link:ort) %>% 
  filter(is.na(strasse) | link %in% problems) %>% 
  mutate(strasse = ifelse(institution == "ALTERS- & PFLEGEHEIM ST. ANTONIUS",
                          "Innere Saastalstrasse 107", strasse)) %>% 
  mutate(strasse = ifelse(institution == "ALTERS- & PFLEGEHEIM SUNNUSCHII",
                          "Dorf Guttet 13", strasse)) %>%   
  mutate(strasse = ifelse(institution == "BETAGTENZENTRUM ÄTTENBÜHL", 
                          "Aettenbühl 2", strasse)) %>% 
  mutate(strasse = ifelse(institution == "ALTERS- & PFLEGEHEIM RÜTTIHUBELBAD", 
                          "Rüttihubel 29", strasse)) %>% 
  mutate(address_orig = if_else(!is.na(strasse),
                                paste(institution, strasse, ort, sep = ", "),
                                paste(institution, ort, sep = ", "))) %>% 
  mutate(address_orig = paste(address_orig, "Switzerland", sep = ", "))
```

Attempting to rescue `r number(nrow(bag_addresses_no_street), big.mark = ",")` addresses with various missing parts of address, mostly street name.

```{r google-proc, eval=FALSE, include=FALSE}
# geocode, reproject to CH1903 / LV03, extract both coords to cols
results_google <- bag_addresses_no_street %>% 
  mutate_geocode(address_orig, output = "more", ext = "ch") %>% 
  rename(address_google = address) %>% 
  select(-north, -south , -east, -west) 

# manual corrections
results_google %<>% 
  mutate(source = "google") %>% 
  rename(address_geocoder = address_google) %>% 
  mutate(lon = ifelse(institution == "ALTERS- & PFLEGEHEIM ST. ANTONIUS", 
                      7.93577374581253, lon)) %>% 
  mutate(lat = ifelse(institution == "ALTERS- & PFLEGEHEIM ST. ANTONIUS", 
                      46.13062054756855, lat)) %>% 
  mutate(source = ifelse(institution == "ALTERS- & PFLEGEHEIM ST. ANTONIUS", 
                         "manual", source)) %>% 
  mutate(lon = ifelse(institution == "ALTERS- & PFLEGEHEIM RÜTTIHUBELBAD", 
                      7.59856495803135, lon)) %>% 
  mutate(lat = ifelse(institution == "ALTERS- & PFLEGEHEIM RÜTTIHUBELBAD", 
                      46.939927400209825, lat)) %>% 
  mutate(source = ifelse(institution == "ALTERS- & PFLEGEHEIM RÜTTIHUBELBAD", 
                         "manual", source)) %>% 
  mutate(lon = ifelse(institution == "ALTERS- & PFLEGEHEIM FLÄCKEMATTE", 
                      8.266983206768096, lon)) %>% 
  mutate(lat = ifelse(institution == "ALTERS- & PFLEGEHEIM FLÄCKEMATTE", 
                      47.09349358397919, lat)) %>% 
  mutate(source = ifelse(institution == "ALTERS- & PFLEGEHEIM FLÄCKEMATTE", 
                         "manual", source)) %>% 
  mutate(lon = ifelse(institution == "ALTERS- & PFLEGEHEIM EMSERBERG", 
                      7.6933695477414235, lon)) %>% 
  mutate(lat = ifelse(institution == "ALTERS- & PFLEGEHEIM EMSERBERG", 
                      46.29313254652122, lat)) %>% 
  mutate(source = ifelse(institution == "ALTERS- & PFLEGEHEIM EMSERBERG", 
                         "manual", source)) %>% 
  mutate(lon = ifelse(institution == "ALTERS- & PFLEGEHEIM SERNFTAL", 
                      9.167769400562054, lon)) %>% 
  mutate(lat = ifelse(institution == "ALTERS- & PFLEGEHEIM SERNFTAL", 
                      46.91622103311342, lat)) %>% 
  mutate(source = ifelse(institution == "ALTERS- & PFLEGEHEIM SERNFTAL", 
                         "manual", source)) %>% 
  mutate(lon = ifelse(institution == "ALTERS- & PFLEGEHEIM SUNNUSCHII", 
                      7.6668983196833365, lon)) %>% 
  mutate(lat = ifelse(institution == "ALTERS- & PFLEGEHEIM SUNNUSCHII", 
                      46.32432057290535, lat)) %>% 
  mutate(source = ifelse(institution == "ALTERS- & PFLEGEHEIM SUNNUSCHII", 
                         "manual", source))

write_rds(results_google, "data/geocoding/results_google.Rds")
```

```{r}
read_rds(results_google, "data/geocoding/results_google.Rds")

results_google_clean <- results_google %>% 
  st_as_sf(coords = c("lon", "lat"), remove = FALSE,
           crs = 4326, 
           agr = "identity") %>%  
  st_transform(crs = 2056) %>% 
  dplyr::mutate(geoy = sf::st_coordinates(.)[,1],
                geox = sf::st_coordinates(.)[,2]) %>% 
  st_drop_geometry() %>% 
  select(link, address_geocoder, geoy, geox, source)

rm(bag_addresses_no_street, results_google)
```

Rescued `r number(nrow(filter(results_google_clean, source == "manual")), big.mark = ",")` addresses with manual searches.


