---
title: "SOMED-geo project"
subtitle: "Linking SOMED to STATPOP"
description: |
  Spatial linkage to building and person file to asses building type and population distribution.
date: "`r Sys.Date()`"
author:
  - name: Radoslaw Panczak 
    url: https://github.com/RPanczak
    affiliation: ISPM
    affiliation_url: https://www.ispm.unibe.ch/
    orcid_id: 0000-0001-5141-683X
# bibliography: biblio.bib
output:
  distill::distill_article:
    highlight: pygments
    toc: true
    toc_depth: 1
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width=9, fig.height=7, dpi=300, out.width="900px", out.height="700px")

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

options(scipen=999)
set.seed(12345)

library(pacman) 
p_load(tidyverse, haven, janitor, magrittr, 
       kableExtra, scales, hrbrthemes,
       sf, nngeo, tmap, tmaptools, geofacet)

tmap_mode("view") # makes map interactive
```

# Data

## SOMED 

```{r}
bag_addresses_clean_geo <- read_rds("data/BAG-open/bag_addresses_clean_geo.Rds")
```

## STATPOP

```{r include=FALSE}
r18_bu_orig <- read_dta("../SNC_core/data-raw/statpop/r18_bu_orig.dta")
# View(r18_bu_orig)

r18_bu <- r18_bu_orig %>% 
  select(-r18_cant, -r18_cant_sh, -r18_gdename, 
         -r18_ecoord_h, -r18_ncoord_h, 
         -r18_plzz,
         -r18_fgkod, -r18_buildper, 
         -r18_nrfloor, 
         -r18_quarter, -r18_comm) %>% 
  # filter(!is.na(r18_ncoord)) %>% 
  # filter(!is.na(r18_ecoord)) %>% 
  st_as_sf(coords = c("r18_ecoord", "r18_ncoord"), 
           crs = 2056,
           remove = TRUE)

sjmisc::frq(r18_bu$r18_buildcat)
sjmisc::frq(r18_bu$r18_gklas)
sjmisc::descr(r18_bu$r18_nrdwell)
```

Using data from `STATPOP` **2018** consisting of `r scales::number(nrow(r18_bu), big.mark = ",")` buildings.

```{r include=FALSE}
rm(r18_bu_orig)
gc()
```

# Link

## Distances 

```{r}
somed_r18_bu <- st_join(bag_addresses_clean_geo, r18_bu, join = st_nearest_feature)

nearest <- st_nearest_feature(bag_addresses_clean_geo, r18_bu)
somed_r18_bu$dist <- st_distance(bag_addresses_clean_geo, r18_bu[nearest, ], 
                                by_element = TRUE)
summary(somed_r18_bu$dist)

# sjPlot::plot_frq(as.numeric(somed_r18_bu$dist), type = "histogram", show.mean = TRUE)

ggplot(somed_r18_bu, aes(as.numeric(dist))) + 
  geom_histogram(binwidth = 10) + 
  theme_minimal() +
  xlab("Distnace to nearest building") + ylab("Number of facilities")
```

## Points in 25 m range 

```{r}
somed_r18_bu_mid <- somed_r18_bu %>% 
  mutate(dist = as.numeric(dist)) %>% 
  filter(dist < 25) %>% 
  select(link:source)

somed_r18_bu_mid_nn <- st_join(somed_r18_bu_mid, r18_bu, join = st_nn, k = 5) 

somed_r18_bu_mid_nn_geo <- r18_bu %>% 
  filter(v0_buildid %in% somed_r18_bu_mid_nn$v0_buildid)

tm_shape(somed_r18_bu_mid) +
  tm_dots(col = "red") +
  tm_shape(somed_r18_bu_mid_nn_geo) +
  tm_dots()
```

# Characteristics

## All included

```{r}
sjmisc::frq(somed_r18_bu$r18_buildcat)
sjmisc::frq(somed_r18_bu$r18_gklas)
summary(somed_r18_bu$r18_nrdwell)
```

## Already defined 

Still few buildings with low (even zero!) number of residents on all three possible measures of pop?

```{r}
select_gklass <- somed_r18_bu %>% 
  filter(r18_gklas > 1122 & !is.na(r18_gklas))

sjmisc::frq(select_gklass$r18_buildcat)
sjmisc::frq(select_gklass$r18_gklas)
summary(select_gklass$r18_nrdwell)
summary(select_gklass$r18_nrpers_total)
summary(select_gklass$r18_nrpers_main)
summary(select_gklass$r18_nrpers_perm)
```

## Missning

There might be some hidden here?

```{r}
select_missing <- somed_r18_bu %>% 
  filter(is.na(r18_gklas))

sjmisc::frq(select_missing$r18_buildcat)

summary(select_missing$r18_nrdwell)
summary(select_missing$r18_nrpers_total)
summary(select_missing$r18_nrpers_main)
summary(select_missing$r18_nrpers_perm)
```

## Residential 

Shouldn't be here at all? Unless with large `dist`?

```{r}
select_done <- somed_r18_bu %>% 
  filter(between(r18_gklas, 1110, 1122)) 

sjmisc::frq(select_done$r18_buildcat)
sjmisc::frq(select_done$r18_gklas)
summary(select_done$r18_nrdwell)
summary(select_done$r18_nrpers_total)
summary(select_done$r18_nrpers_main)
summary(select_done$r18_nrpers_perm)

summary(select_done$dist)
```
